---
title: "R Notebook"
output: html_notebook
---

First, from the 19 climatic variables of chelsa (30 seconds), I crop the extension to Asia (62.86653, 113.9249, 1.566527, 37.66653) (xmin, xmax, ymin, ymax). See previous script of how I did this.

However, this is still a super large file. Let's resample the bio files in a 10-fold order. dimensions : 4332, 6127, 26542164  (nrow, ncol, ncell).
I converted to the ascii format because it's the only format Maxent accepts.

```{r}
library(raster)

#Asia file names
asia_tifs = Sys.glob('asia_bio_*.tif')

#Loop to read the files as rasters, resample the original rasters and then write
#them as new raster files in ascii format

for(i in 1:length(asia_tifs)){
 bio_r = raster(asia_tifs[i])
 bio_aggr = aggregate(bio_r, fact=10)
 writeRaster(bio_aggr, filename=paste0(names(bio_r), '_resampled.asc'), format = 'ascii')
}

```

With these new resampled raster, let's conduct some correlation analyses to remove high-correlated variables.


```{r}
library(caret)
library(corrplot)

#Read the new resampled files
asia_ascs = Sys.glob('asia_*resampled.asc')

#stack them as rasters
asia_stack = stack(asia_ascs)

#extract the values of these raster files
asia_values = values(asia_stack)

#remove NAs
asia_values = na.omit(asia_values)

#Create a correlation matrix
asia_corr = cor(asia_values)

#setting a threshold of 70%
cor_threshold = 0.7

#Check which one of the variables are highly correlated
highly_correlated = as.vector(findCorrelation(asia_corr, cutoff = cor_threshold))
highly_correlated_names = names(asia_stack)[highly_correlated]

#and which are not that correlated (i.e., those that we are most interested)
less_correlated = which(names(asia_stack) %in% setdiff(names(asia_stack), highly_correlated_names))

less_correlated_names = setdiff(names(asia_stack), highly_correlated_names)

corrplot(asia_corr, type = "lower", diag = F, tl.srt = 45, mar = c(3, 0.5, 2, 1),
         title = "Correlation between variables of CHELSA bioclimatic variables")



```

So, we'll keep only 8 variables:

[1] "asia_bio_10_resampled" "asia_bio_13_resampled" "asia_bio_14_resampled"
[4] "asia_bio_15_resampled" "asia_bio_18_resampled" "asia_bio_19_resampled"
[7] "asia_bio_2_resampled"  "asia_bio_3_resampled" 

Now, let's conduct some Niche Modelling using Maxent.
I'll conduct these niche analyses per lineage (Curcuma, Curcuma I, Curcuma II,
Ecomata, Hitcheniopsis, Pierrana and Roscoeana).

After conducting the Niche MOdelling using Maxent, we'll extract the values of the 8 variables following the the shortest distance to the top-left corner (0,1) in ROC plot (check the html files for the AUC value).
```{r}
library(raster)

#the less correlated variables
asia_eight_stack = stack(paste0('./env_variables/', less_correlated_names, '.asc'))

#our maxent niche modelling
cs_model = raster('./maxent_output/curcuma_cs.asc')
c1_model = raster('./maxent_output/curcuma_cs1.asc')
c2_model = raster('./maxent_output/curcuma_cs2.asc')
e_model = raster('./maxent_output/curcuma_es.asc')
h_model = raster('./maxent_output/curcuma_hs.asc')
p_model = raster('./maxent_output/curcuma_ps.asc')
r_model = raster('./maxent_output/curcuma_rs.asc')

#get the pixels where the niche analysis was more or equal than the top-left cornern in ROC plot.
cs_model_89 = cs_model
c1_model_88 = c1_model
c2_model_92 = c2_model
e_model_94 = e_model
h_model_94 = h_model
p_model_94 = p_model
r_model_94 = r_model

cs_model_89[] = ifelse(values(cs_model) >= 0.89, 1, NA)
c1_model_88[] = ifelse(values(c1_model) >= 0.88, 1, NA)
c2_model_92[] = ifelse(values(c2_model) >= 0.92, 1, NA)
e_model_94[] = ifelse(values(e_model) >= 0.94, 1, NA)
h_model_94[] = ifelse(values(h_model) >= 0.94, 1, NA)
p_model_94[] = ifelse(values(p_model) >= 0.94, 1, NA)
r_model_94[] = ifelse(values(r_model) >= 0.94, 1, NA)

#select the values from bioclimatic variables that matches where the niche 
#modeling was >= the top-left cornern in ROC plot.

asia_masked_by_cs_model = mask(asia_eight_stack, cs_model_89)
asia_masked_by_cs_model_df = as.data.frame(rasterToPoints(asia_masked_by_cs_model))

asia_masked_by_c1_model = mask(asia_eight_stack, c1_model_88)
asia_masked_by_c1_model_df = as.data.frame(rasterToPoints(asia_masked_by_c1_model))

asia_masked_by_c2_model = mask(asia_eight_stack, c2_model_92)
asia_masked_by_c2_model_df = as.data.frame(rasterToPoints(asia_masked_by_c2_model))

asia_masked_by_e_model = mask(asia_eight_stack, e_model_94)
asia_masked_by_e_model_df = as.data.frame(rasterToPoints(asia_masked_by_e_model))

asia_masked_by_h_model = mask(asia_eight_stack, h_model_94)
asia_masked_by_h_model_df = as.data.frame(rasterToPoints(asia_masked_by_h_model))

asia_masked_by_p_model = mask(asia_eight_stack, p_model_94)
asia_masked_by_p_model_df = as.data.frame(rasterToPoints(asia_masked_by_p_model))

asia_masked_by_r_model = mask(asia_eight_stack, r_model_94)
asia_masked_by_r_model_df = as.data.frame(rasterToPoints(asia_masked_by_r_model))

#create a new column (for ecospat analysis) setting one for where the niche is 
# >= the top-left cornern in ROC plot.
asia_masked_by_cs_model_df$species_occ = 1
asia_masked_by_c1_model_df$species_occ = 1
asia_masked_by_c2_model_df$species_occ = 1
asia_masked_by_e_model_df$species_occ = 1
asia_masked_by_h_model_df$species_occ = 1
asia_masked_by_p_model_df$species_occ = 1
asia_masked_by_r_model_df$species_occ = 1

#Now, select the values from bioclimatic variables that doesn't match to our
#threshould (>= the top-left cornern in ROC plot)
asia_masked_by_cs_invert = mask(asia_eight_stack, cs_model_89, inverse = TRUE)
asia_masked_by_c1_invert = mask(asia_eight_stack, c1_model_88, inverse = TRUE)
asia_masked_by_c2_invert = mask(asia_eight_stack, c2_model_92, inverse = TRUE)
asia_masked_by_e_invert = mask(asia_eight_stack, e_model_94, inverse = TRUE)
asia_masked_by_h_invert = mask(asia_eight_stack, h_model_94, inverse = TRUE)
asia_masked_by_p_invert = mask(asia_eight_stack, p_model_94, inverse = TRUE)
asia_masked_by_r_invert = mask(asia_eight_stack, r_model_94, inverse = TRUE)


asia_masked_by_cs_invert_df = as.data.frame(rasterToPoints(asia_masked_by_cs_invert))

asia_masked_by_c1_invert_df = as.data.frame(rasterToPoints(asia_masked_by_c1_invert))

asia_masked_by_c2_invert_df = as.data.frame(rasterToPoints(asia_masked_by_c2_invert))

asia_masked_by_e_invert_df = as.data.frame(rasterToPoints(asia_masked_by_e_invert))

asia_masked_by_h_invert_df = as.data.frame(rasterToPoints(asia_masked_by_h_invert))

asia_masked_by_p_invert_df = as.data.frame(rasterToPoints(asia_masked_by_p_invert))

asia_masked_by_r_invert_df = as.data.frame(rasterToPoints(asia_masked_by_r_invert))

#create a new column (for ecospat analysis) setting zero for where the niche is 
# < the top-left cornern in ROC plot.
asia_masked_by_cs_invert_df$species_occ = 0
asia_masked_by_c1_invert_df$species_occ = 0
asia_masked_by_c2_invert_df$species_occ = 0
asia_masked_by_e_invert_df$species_occ = 0
asia_masked_by_h_invert_df$species_occ = 0
asia_masked_by_p_invert_df$species_occ = 0
asia_masked_by_r_invert_df$species_occ = 0

#bind together the two dfs (where is >= the top-left cornern in ROC plot and where is < the top-left cornern in ROC plot)
cs_to_ecospat = rbind(asia_masked_by_cs_model_df, asia_masked_by_cs_invert_df)
c1_to_ecospat = rbind(asia_masked_by_c1_model_df, asia_masked_by_c1_invert_df)
c2_to_ecospat = rbind(asia_masked_by_c2_model_df, asia_masked_by_c2_invert_df)
e_to_ecospat = rbind(asia_masked_by_e_model_df, asia_masked_by_e_invert_df)
h_to_ecospat = rbind(asia_masked_by_h_model_df, asia_masked_by_h_invert_df)
p_to_ecospat = rbind(asia_masked_by_p_model_df, asia_masked_by_p_invert_df)
r_to_ecospat = rbind(asia_masked_by_r_model_df, asia_masked_by_r_invert_df)

write.csv(cs_to_ecospat, 'cs_to_ecospat.csv', row.names = F)
write.csv(c1_to_ecospat, 'c1_to_ecospat.csv', row.names = F)
write.csv(c2_to_ecospat, 'c2_to_ecospat.csv', row.names = F)
write.csv(e_to_ecospat, 'e_to_ecospat.csv', row.names = F)
write.csv(h_to_ecospat, 'h_to_ecospat.csv', row.names = F)
write.csv(p_to_ecospat, 'p_to_ecospat.csv', row.names = F)
write.csv(r_to_ecospat, 'r_to_ecospat.csv', row.names = F)
```


